### 2019-02-06 sections/structure.tex

% !TEX root=../icfp2019.tex

% \paragraph{Overview}

In Section~\ref{sec:example} we demonstrate the functionality of \TOPHAT by means of an example,
Section~\ref{sec:comparison} compares \TOP and \TOPHAT to \CSP.
Section~\ref{sec:language} introduces the \TOPHAT calculus syntax
and Section~\ref{sec:semantics} the semantics.
Then in Section~\ref{sec:properties} we show that certain properties hold for the calculus.
We take a look at related work in Section~\ref{sec:relatedwork}
and conclude in Section~\ref{sec:conclusions}.