### 2019-02-19 sections/outlook.tex
rename future work to outlook

% !TEX root=../icfp2019.tex

\section{Outlook}

\fixme{write this section}