### 2019-02-19 sections/introduction.tex
Fixed contributions in introduction

% !TEX root=../icfp2019.tex

\section{Introduction}

\fixme{Common intro??}

\subsection{Tasks}

Many applications these days are developed to support workflows in institutions and businesses.
Take for example expense declarations, order processing, and emergency management.
Some of these workflows occur on the boundary between organisations and customers,
like flight bookings or tax returns.
What they all have in common is
that they need to interact with different people (end users, tax officers, customers, etc.)
and they use information from multiple sources (input forms, databases, sensors, etc.).

We call interactive units of work based on information sources \emph{tasks}.
% Tasks stand for units of work in the real world, assigned to people.
Tasks model real world collaboration between users,
are driven by work users do,
and are assigned to some user.
Users could be people out in the field or sitting behind their desks,
as well as machines doing calculations or fetching data.

\subsection{Task-oriented programming}

% Task-Oriented Programming (\TOP) is a programming paradigm aimed at writing interactive multi-user applications in a declarative way \cite{conf/ppdp/PlasmeijerLMAK12}.
Task-oriented programming (\TOP) is a programming paradigm which aims to be the sweet spot between faithful modelling workflows
and rapid prototyping of multi-user web applications supporting these workflows.
% Task-oriented programming (\TOP) is a programming paradigm to support these ways of working.
\TOP focusses on modelling collaboration patterns.
This gives rise to user's need to interact and share information.
Next to that, \TOP automatically provides solutions to common development jobs like designing \GUI\ s, connecting to databases, and communicating between servers and clients.

Therefore,
a language that supports \TOP should choose the right level of abstraction to support two things.
% \TOP has two aspects.
% First, it should allow to specify tasks from real world scenarios.
Firstly, it should provide primitive building blocks that are useful for high-level descriptions of how users collaborate with each other and with machines.
These building blocks are \emph{editors}, \emph{composition}, and \emph{shared data}.
% Second, it should be able to generate multi-user web applications to support these scenarios.
Secondly, it should be able to generate applications, including graphical user interfaces, from workflows modelled with said building blocks.

Users can work together in a number of ways, and this is reflected in \TOP by task compositions.
There is sequential composition, parallel composition, and choice.
Users need to communicate in order to engage in these forms of collaboration.
This is reflected in \TOP by three kinds of communication mechanisms.
There is data flow \emph{alongside} control flow, where the result of a task is passed onto the next.
There is data flow \emph{across} control flow, where information is shared between multiple tasks.
Finally, there is communication with the \emph{outside} world, where information is entered into the system via input events.
The end points where the outside world interacts with \TOP applications are called editors.
In generated applications, editors can take many forms, like input fields, selection boxes, or map widgets.

\fixme{Change if title changes.}
In this paper we study the essence of \TOP.
We claim that at the heart of \TOP lie modular interactive workflows.
Workflows because \TOP programs coordinate collaboration between people.
Interactive because progress of \TOP programs is driven by user input.
Modular because the embedding in a strongly-typed functional language allows powerful abstractions using type-driven techniques and higher-order functions.

\subsection{Utilisation}

Currently, we know of two frameworks using \TOP: \ITASKS and \MTASKS.

\ITASKS is an implementation of \TOP, in the form of a shallowly embedded domain-specific language in the lazy functional programming language Clean.
It is a library that provides editors, monadic combinators, and shared data sources.
\ITASKS uses the generic programming facilities of Clean to derive rich client and server applications from a single source.
It has been used to model an incident management tool for the Dutch coast guard~\cite{conf/iscram/LijnseJP12}.
Also it has been used numerous times to prototype ideas for Command and Control~\cite{theses/nlda/Kool17, theses/radboud/Stutterheim17}, and in a case study for the Dutch tax authority \cite{conf/sfp/StutterheimAP17}.

\MTASKS is a subset of \ITASKS,
focussing on \IOT devices and deployment on micro controllers.
\fixme{Add examples.}

\fixme{Say what lacks in these approaches: no formal reasoning about properties.}

\subsection{Scope}

In this paper, we set out to develop a formal basis for \TOP, called \TOPHAT (TopHat).
Our goal is to specify the basic building blocks of declarative workflows.
We clearly separate them from general purpose programming concepts.
This allows us to prove properties like progress and type preservation.
\TOPHAT should be powerful enough to model complicated real-world situations,
without losing opportunities for formal reasoning.

By separating them in the style of a deep embedding \cite{conf/cefp/Gibbons13}, we make explicit where and where not, they depend on each other. \fixme{what to do with this sentence}

We show that, where process algebras focus on the \emph{coordination} of processes and the \emph{communication} between them,
\TOP, and therefore \TOPHAT, focus on supporting \emph{tasks} and model ways to \emph{collaborate} on them.

In this paper we aim to identify the essence of \TOP in formal as well as an informal way.
These core principles lead to a formal system for \TOP called \TOPHAT.
\TOPHAT paves the way to formal reasoning about applications written using the \TOP paradigm,
like those in \ITASKS and \MTASKS.

\TOP applications give rise to concepts like concurrency, synchronisation, \GUI-programming and program generation.
In this paper we identify common concepts from already existing languages and frameworks,
and describe their similarities and differences.\fixme{What do you mean with this sentence?}

\subsection{Contributions}

Our contributions to workflow modelling, functional programming language design, and rapid application development are as follows.

\begin{itemize}
  \item
    We informally describe the essential concepts of task-oriented programming (\TOP).

  \item
    We present a formal calculus for modelling declarative workflows, embedded in a simply typed lambda-calculus.
    This calculus is based on the aforementioned essential \TOP concepts.

  \item
    We develop an operational semantics that is driven by user input.
    The semantics of the task language is clearly separated from the semantics of the underlying host language.

  \item
    Along with the semantics, we present the following semantic observations on tasks:
    the current value, whether a term is stuck, the current user interface, and the accepted inputs. \fixme{we do not provide observation for user interface!}

  \item
    We implemented the whole semantic system in the dependently typed programming language Idris \cite{journals/jfp/Brady13}.

  \item
    We have implemented \TOPHAT in iTasks to demonstrate that the first is a subset of the second.

  \item
    Using both informal and formal descriptions, we compare \TOP with multiple related works in areas,
    ranging from business process modelling languages, to process algebras and reactive programming frameworks.

    \item
      We prove progress and type preservation for our calculus.

\end{itemize}

\subsection{Structure}

In Section~\ref{sec:example} we demonstrate the functionality of \TOPHAT by means of an example,
Section~\ref{sec:intuition} gives an overview of the essential concepts of \TOP and \TOPHAT.
Section~\ref{sec:language} introduces the \TOPHAT calculus syntax
and Section~\ref{sec:semantics} the semantics.
Then in Section~\ref{sec:properties} we show that certain properties hold for the calculus.
We take a look at related work in Section~\ref{sec:relatedwork}
and conclude in Section~\ref{sec:conclusions}.