### 2019-02-20 sections/outlook.tex
Initial setup outlook section

% !TEX root=../icfp2019.tex

\section{Outlook}

The work presented in this paper describes a formal foundation for task oriented programming.
For the first time, this programming paradigm has been formalised.
The formalisation allows us to reason about \TOP and programs written in \TOPHAT.
So far, we have proven type safety and progress, but this is merely the beginning.
In future work, we plan on utilising the formalisation to do much more.

First of all, we would like to show that certain properties hold for individual programs.
We would also like to be able to define what it means for programs to be equal.

A more in depth description of future work can be found in Section~\ref{sec:conclusions}.