### 2019-04-18 reviews/icfp-before-rebuttal.txt
add icfp reviews

ICFP 2019 Paper #111 Reviews and Comments
===========================================================================
Paper #111 TopHat: A formal foundation for task-oriented programming

Review #111A
===========================================================================

Overall merit
-------------
B. OK paper, but I will not champion it.

Reviewer expertise
------------------
Y. I am knowledgeable in the area, though not an expert.

Paper summary
-------------
Interactive workflow "tasks" are a common use of computing
resources. This paper provides a middleweight calculus modeling key
ideas of such interactions: a set of combinators specifies the inputs,
computations, steps, and orderings in such tasks.

The model is a subset of features of existing implementations iTasks
and mTasks, though there's no formal relationship.

The semantics are proven type safe and a small model is implemented in
Idris (though no detail is given on this mechanized model).

The paper concludes with a substantial comparison to related work.

Comments for author
-------------------
The work here is a nice semantics on an important topic that has
received too little attention in the PL community. My score is a B for
several reasons:

- Nothing is _done_ with the semantics beyond sanity checking. Such
  type safety proofs show only that you've built a sensible
  system. What is the semantics for?

- The papers aims and claims don't match up with its actual
  contributions.

  While "formal program verification is important for mission-critical
  software", no actual formal verification happens here---and there's
  no evidence that the workflows described here are mission critical.

  The "_____-oriented programming" framing leads one to compare the
  system to other orientations, which may be unfavorable either
  because other orientations are total views of what programming is
  (and so more substantial ideas than what any single semantics might
  offer) or because other orientations are seen as software
  engineering fads.

- The semantics is entirely operational, which offers the weakest
  reasoning principles as compared with denotational or axiomatic
  models. The semantics and combinators also have gaps, allowing one
  to express senseless workflows.

Despite these flaws, I think the combinators given here will be useful
elsewhere, so I lean towards acceptance.

# What might improve the paper?

- Moderate the claims. The key novelty here is an interesting library
  of combinators in an important domain and a nice semantics for a
  non-recursive fragment. Your properties don't quite "validate" your
  semantics (line 838), but rather serve as a sanity check.

- Use the semantics for something. An Iris-like logic would more than
  suffice for proofs of equivalence of programs. Why not prove that
  one can always eventually eat breakfast? (Such a property is highly
  desirable and sure to be popular.)

  In the related work, you frequently remark that other systems don't
  support formal reasoning---such statements would be more sensible if
  you used your semantics for formal reasoning.

- Summarize the interaction trees of tasks. Throughout the paper, it's
  stressed that TopHat merely specifies collaboration rather than
  explicit communication, as in, e.g., process calculi or session
  types. While such a system is perhaps nice to program in, having a
  term `e : Task t` doesn't tell me much of anything about the set of
  input events `e` can receive. Regular expressions (with embedded
  functional programs to enable transitions) should suffice here, and
  will make for a nice, abstract denotational semantics of tasks.

- Prove that your combinators are orthogonal. Pierce's spider calculus
  (2010) is a nice example: each primitive is proved 'independent' of
  the others.

- Clarify which kinds of tasks are realizable. There's no monadic
  "join" operation that allows us to work with `Task (Task t)`, yet
  the type is expressible. The set of editors is also surely limited:
  what can one do with a $\square (\lambda x:\textsc{Int}. x)$? What
  about $\blacksquare (\textsc{Int}\rightarrow\textsc{Int})$? What
  about $\square (\square 5)$? What about $\blacksquare (\textsc{Task}
  ~ \textsc{Int})$?

  Relatedly, tasks are values in the host language (line 591). What
  eliminators on tasks do you offer?

- You are able to write the smoker program simply because your
  semantics eats all the complexity. What would dining philosophers
  look like in your system? It seems you cannot write a version of it
  that dynamically determines the number of philosophers. What might
  you need to add to the language to support that?

  Later on (line 472) you talk about "immediately reflected" changes,
  but consistent, available (i.e., immediate), and persistent
  propagation is impossible in a distributed system. How should we
  think about the possibility of network delay or failure?
  Communication-based systems force programmers to think about these
  issues, for better or worse.

  Similarly, your comparison to HipHop misses a crucial fact: HipHop's
  Esterel-like semantics has no data races, while TopHat is full of
  data races. While not themselves a problem, there's strong anecdotal
  evidence that racy models are harder to program in.

- Clarify your implementations. What does the Idris implementation do?
  Is it actually executable, or is it a relational model? How does
  your Idris implementation compare to your embedding in iTasks? Have
  you tried testing whether the two semantics behave the same?

- Clarify your relation to FRP. You say that TopHat has no notion of
  time, but $\blacksquare \mathsf{time}$ seems to exactly be a
  time-varying value. It seems that one could use $\blacksquare$ and
  $\RHD$ to write FRP-ish programs, e.g., your derived `delay`
  combinator.

# Questions

You say you "show that the normalisation semantics is a big-step
semantics". Which property is that, and why is it important? Why does
having state complicate a big-step semantics?

Could you encode BPEL in TopHat? If not, which features does BPEL
support that you do not? (E.g., cycles?)

# Minor stuff

- Line 235: What is an invalid event?

- Line 269: What about explicit option types?

- Line 349: We haven't seen these internal operators yet. It might be
  good to at least name them all up front. Could you extend the
  airline example to include them?

- Line 403: You call $\bowtie$ "combination" but perhaps 'parallel
  composition' or 'interleaving' would be more helpful.

- Line 657: It may be worth clarifying that $\rhd$ doesn't need to
  look at $e_2$ because a `C` event is needed. Why doesn't
  $\mathcal{F}$ look at $e_2$ for $\RHD$, though, which needs no
  continue event? Isn't $\blacksquare 5 \RHD \Lightning$ failing?

- Line 793: The typing side conditions on operational are a strange
  violation of the phase distinction.

- Line 949: If iTasks "implicitly has a formal semantics", are there
  any languages that _don't_?

- A few typos: focussing, 'mad an implementation', moddeling

Review #111B
===========================================================================

Overall merit
-------------
B. OK paper, but I will not champion it.

Reviewer expertise
------------------
Z. I am not an expert. My evaluation is that of an informed outsider.

Paper summary
-------------
The paper presents a formalization of the essence of an existing "real-world task" workflow modelling language and methodology called TOP

Comments for author
-------------------
I am an outsider for this particular area of programming languages and was not familiar with TOP.

The paper is well written and interesting, especially for those who know and use TOP and need a detailed description of its formal semantics and some of its properties.

If I have misgivings, I think it is about TOP itself - the question is not its soundness, or its semantics, but its utility for the business domain and the "logical encoding" approach used to formalizing real-world business systems.  My own experience has been with systems such as Oracle Policy Automation which approaches modelling of real-world processes in a vastly different way.  So the sea of symbols as an "encoding" of the real world is imply daunting and unfamiliar to me, and I am skeptical of its utility.  However, since the focus of the paper is on formalization and not on TOP itself, this is me just admitting a (perhaps unfair) bias.

I've given my review a low confidence rating as a result.

Review #111C
===========================================================================

Overall merit
-------------
C. Weak paper, though I will not fight strongly against it.

Reviewer expertise
------------------
Y. I am knowledgeable in the area, though not an expert.

Paper summary
-------------
The paper presents TopHat, an imperative higher-order language with facilities
for interacting with an external environment.

In typical use cases, the environment can be thought of as a user interacting
with the program through a graphical interface. TopHat allows a program to
communicate with the environment via "tasks", which are logical descriptions of
input/output interfaces. Compound tasks can be formed via dedicated combinators:
pairing, monadic-bind-like sequential composition, choice, etc.

Technically, the paper contains an introduction to "task-oriented" programming
in TopHat, a formal semantics for TopHat, a type safety result, and a discussion
of related work.

Comments for author
-------------------

I welcome every effort to make UI programming easier, given the sorry state of
current programming practice. As such, I am largely sympathetic towards the aims
of "task-oriented programming", and find this paper intriguing. Unfortunately,
I'm not convinced that it would be a good fit for ICFP in its current form.

First, I don't think the paper contains enough technical contributions. I
understand that its goal is to give a standalone formalization of task-oriented
programming, which has been around for some time but always in the form of
libraries. But as is, this formalization does not pay off: all applications are
left for future work, which I find highly unsatisfactory. Apparently, we did not
learn anything new about the equivalence of task-oriented programs, nor gain
insight into their implementation, etc. Such points are left for future work,
but would greatly improve the relevance of the work.

Second, I am doubtful of certain design aspects of TopHat.

For example, the language mixes a form of shared memory with bespoke concurrency
constructs, which has several important consequences. In particular, sequential
composition performs speculative execution: if t is a task with a set value, the
sequential composition combinator t ▷ e tries to execute e with this value, and
backtracks if e fails ("failing" is an explicit concept in TopHat),
reestablishing memory as if e had never been executed. This makes composition
very expressive (see e.g. the example given in §3.6, p. 7), but also very
costly, as, in the worst case, the entire store must be duplicated. But I am
even more worried about the impact of this semantics on reasoning: can one
really understand what's going on in a TopHat program which is not very small?
Doesn't this make it hard to understand the complexity of TopHat programs?

Also, while TopHat seems to be a language dedicated to reactive programming,
i.e., continuous interaction with an environment, it does not help with managing
this interaction in a modular way. As far as I see, there is no way for a task A
to interpose itself between the environment and another task B, i.e., for A to
encapsulate B in a way that is transparent from B's perspective. This is
possible in the reactive languages I know of (e.g., Esterel).

Third, in my opinion the paper is poorly written and hard to understand. It
contains too many forward references and makes use of symbols or notions before
they are defined (see below). I also found several statements hard to interpret,
e.g. the task pairing combinator supposedly "introduces interleaving
concurrency" (p. 7), but as far as I see the semantics of TopHat is perfectly
deterministic (given a fixed input sequence) and does not involve interleaved
access to the store. My confusion was exacerbated by the related work section,
where TopHat is compared in vague terms to works whose relevance is not entirely
obvious to me (e.g., process algebra) but where closely-related work from the
ICFP community, such as Carlsson and Hallgren's Fudgets or Conal Elliott's
Tangible Values, is not mentioned at all. Similarly, the comparison to reactive
programming works are not technical enough in my opinion.

## Detailed comments

p. 4

- The "lightning strike" symbol used in the examples should be introduced in the
  text as soon as it is used.

- The "bookFlight ⋈ bookFlight ⋈ bookFlight" example is very problematic. At
  this point, I had zero understanding of TopHat's concurrency model, and thus
  this program looked like it had a data race, making it plainly incorrect. As a
  consequence, I found the paragraphs extolling the virtues of shared-memory
  programming flabbergasting! This should be addressed immediately in the text,
  explaining what is atomic and what isn't.

p. 5

- While discussing the semantics of parallel composition (⋈), it is stated that
  "what is meant by parallel is that the order in which we do tasks and the
  smaller tasks that they are composed of, does not matter". But I cannot
  reconcile this statement with the formal semantics: ⋈ is not commutative at
  all, since for instance "l := True ⋈ l := False" will always result in a store
  containing l ↦ True, and "l := False ⋈ l := True" in one containing l ↦ False.

- "Inputs are synchronous."

  I believe that only a single task may evolve per time step, judging from the
  grammar of inputs presented in Fig 14. So what does this mean, exaclty? That
  program and environment alternate?

p. 6

- I would have liked an intuitive description of the graphical representation of
  choice combinators.

p. 7

- "The details of the solution [Downey's] are not important here."

  Then perhaps one could simplify the preceding paragraph?

p. 8

- Internal choice is mentioned here, without ever being introduced before!

p. 9

- The grammar allows function-valued editors. What does ⊠(Int → Int) look like?

- The grammar allows higher-order tasks and editors. What does ⊠⊠Int look like?

p. 10

- "If e is a reference of type Ref τ, then ▪e is of type Task τ"

  Here τ should be changed into β, or T-Update should be modified.

- The rationale for the external vs. internal distinction is never given. I
  would expect that in an expressive language, external combinators would be
  definable in terms of their internal counterparts.

p. 11

- I liked example 4.4.

p. 13

- "When both branches have a value, it takes the value of the left branch"

  Why not give some combination function?

p. 14

- In Figure 10, F is a function only if ⇓ is deterministic. I believe it is, but
  perhaps this should be stated explicitly somewhere.

p. 15

- Isn't rule S-Xor lacking some premises?

p. 17

- l. 832: obfuscating the link to the public GitHub repository is worse than
  useless. Why not attach an anonymized copy of the source code to the paper?

p. 18

- "we mad": we made.

- "Also, we mad[e] an implementation of TopHat combinators into iTasks. This
  shows that TopHat is a subset of iTasks."

  I'm nitpicking, but I don't like that turn of phrase. Without additional
  information about the simplicity of this implementation, this conclusion is
  incorrect. Otherwise, the fact that you can implement, say, Clean on top of
  x86 assembly would imply that Clean is a subset of x86 assembly.

- l. 926: "understand" -> understanding.

- "We do not know yet if the more complex combinators of iTasks are expressible
  in the basic combinators of TopHat."

  I would have liked to know what these combinators look like.

p. 20

- "However, on IOT devices it is useful to continue running tasks endlessly,
  which is done in mTasks using a "forever" combinator. This is currently not
  possible in TopHat."

  So in TopHat tasks do not "run endlessly", yet they "never terminate" and
  "always keep reacting to events" (p. 6)? This is very confusing.

p. 21

- As I said above, I'm not sure TopHat, a deterministic functional language,
  compares to process algebra.

- l. 996-999: this paragraph is extremely vague ("model collaboration patterns",
  "declarative specification"). Moreover, I don't think comparing "goals" of
  formalisms leads to useful distinctions, otherwise we would conclude that
  functional languages (whose goals are to program computers) and Alonzo
  Church's λ-calculus (whose original goal was to give a foundation for
  mathematics) have nothing to do with each other.

- l. 1015: "syntonisation" -> synchronisation.

- l. 1025: "programatically" -> "programmatically".

- In my opinion, comparing TopHat to ReactiveML would be more relevant, since a
  lot of the HipHop work is about coping with the vagaries of the web
  programming environment.

p. 22

- "FRP and TOP are different systems and have different goals in mind."

  Same remarks as above.

- "TOP has no notion of time" Really? If so, why talk about synchrony and the
  instantaneous character of reactions earlier?

p. 23

- l. 1110: "moddeling" -> "modeling".

Review #111D
===========================================================================

Overall merit
-------------
C. Weak paper, though I will not fight strongly against it.

Reviewer expertise
------------------
Y. I am knowledgeable in the area, though not an expert.

Paper summary
-------------
The paper a formal operational model for a core task-oriented language
inspired by iTasks. The model has a stratified structure. At the bottom
layer there is a base functional typed language with first-class
cells. On top of that there is a language of typed task-related effects.
The semantics of the model is also stratified. At the bottom layer there
is a small-step operational semantics for the base language and on top of
that there are two layers for the effect layer. The bottom of these two
layers deals with task reductions that do not require user input. The
top deals with task reductions that are triggered by user input.
The model is proven type sound and has been implemented in Idris and the
task effect language on top of iTasks.

Comments for author
-------------------
+++++ Comments +++++

++ I think the goal of formalizing the core of iTasks is great. TOP is an
interesting programming paradigm and it would be good to be able to
understand its essence and furthermore to have techniques to prove the
properties of TOP programs.

++ The development of the model seems sensible and the provided examples
are quite interesting.

-- However I think this current paper is far from this goal. As is it
feels more like an exercise in formalisms rather than a research result.
In particular, the approach to describing TOP is a straightforward
adaptation of ideas from process calculi (see work on languages with
session types).

-- Moreover, the task layer and its type system are straightforward
adaptation of monadic effect descriptions. This is a missed opportunity as
a richer type system could be used to prove interesting properties such as
for instance that certain tasks never fail.

-- Finally, there is nothing surprising or challenging in proving type
soundness for the model.

-- This model doesn't meet the requirements to be called a calculus.
Plotkin's seminal paper on the call-by-name and call-by-value lambda
calculus provides a blueprint for how one constructs a calculus for an
existing language. For a more recent example see Florence et al. 's POPL
paper on a calculus for Esterel with the same motivation as this paper but
for Esterel. This effort builds on equatinal theories for effectual programs developed by Felleisen in the 80s. In general, a calculus should allow sound  and consistent
equational reasoning with respect to the existing semantics  of the language
(in this case the graph rewriting one of Clean) and this is possible even in the
presence of effects at the price of weakening reasoning power.

-- Again related to Esterel and the above, the paper ``An axiomatic semantics for
Esterel'' by Simone Tini presents an alternative approach to formalizing
Esterel with the goal of reasoning about Esterel programs. That paper opts
for bisimulation techniques instead of algebraic manipulations of syntax.
Thus similar to TOPHAT, the paper opts for an LTS approach. Again though,
and unlike TOPHAT, the design of the model is directly connected with certain
properties and techniques.

+++++ Small points/typos +++++

page 13: I is mentioned here but it is defined in page 16.

page 18, first line: "we mad" ---> "we made"

page 18, first line: "combinators into" ---> "combinators on top of"

page 21, section 7.3, last line: "syntonisation" ---> "synchronization"

++++++ QUESTIONS THAT THE AUTHORS ARE EXPECTED TO ANSWER ++++++

1) What is challenging/surprising about constructing TOPHAT?

2) Could you please explain what kind of properties of iTasks
programs you intend to intend to try to prove with TOPHAT and
how its design helps this process?